---
title: "WideToLongTableau_INL_112320"
author: "Min Hwangbo"
date: "11/23/2020"
output:
  html_document:
    preserve_yaml: true
    toc: true
    toc_float: true
    keep_md: true
published: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Loading packages - Data reshape
```{r, warning = F, message = F, results = "hide"}
# Install packages first using install.packages("pacakgename") on your console!
## Reference: http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/ 
### Inspiration: Dr. Liz Sanders's HLM class
#### Inspiration II: Chuck Lanfear Intro to R: https://clanfear.github.io/CSSS508/

library("dplyr")  # Data reshaping package
library("tidyr")  # Data transformation package
library("readr")  # CSV loading package
```


# Step 2: Loading data
```{r, msg = F, warning = F}
data.wide <- read_csv("STARS_PLC_20-21.csv") # Loading data from an excel file

# Quality check descriptive
head(data.wide, 5) # First five records
tail(data.wide, 5) # Last five records
ls(data.wide)   # Vars names
## Data summary: 25 records / 37 vars
```

# Step 3a: Changing var names
* First, we're going to copy information on the 2nd row to the 1st row as Var Names.
* Second, we're going to delete the duplicate record.
* Last, we will check the data quality/code accuracy.

```{r}
names(data.wide) <- data.wide[1,] # Copy 2nd row as var names
data.wide_1 <- data.wide[-c(1, 2),] # Delete duplicate rows (1st and 2nd)

head(data.wide_1, 5) # First 5 records
tail(data.wide_1, 5) # Last 5 records
ls(data.wide_1) # Vars name
# 23 obs & 37 vars
```

# Step 3b: Checking `Null` values
```{r}
# summary(data.wide) 
## Data summary: 23 records / 37 vars - check for null values
sum(is.na(data.wide_1$`Which training did you participate in?`))
# 0 NULL RECORDS
```

# Step 3c: Checking finished survey (`FILTER` responses that are considered `finished`)
```{r}
data.wide_2 <- data.wide_1 %>%
  filter(Finished == "True") # Watch out for "True" spelling

# Quality check 
head(data.wide_2, 5) # First five records
tail(data.wide_2, 5) # Last five records
ls(data.wide_2)   # Vars names
# ALl remains the same: 23 obs, 37 vars
```

# Step 4: Data reshape (Wide -> Long)
* Using `gather` function to consolidate questions into one column. 
```{r}
data.long <- gather(data.wide_2, question, response, "Q15 - Content provided matched the training description.":"What changes would you suggest to the trainer?", factor_key = T)

data.long # Check the data
head(data.long, 5) # First five
tail(data.long, 5) # Last five
ls(data.long)
```

# Step 5: Saving as a CSV file
* FileName_ProjectName_MMDDYY(DateCreated)
```{r}
# CSV
write.csv(data.long, "V2STARS_PLC_20-21.csv")
```