---
title: "DataReshape_INL_071320"
author: "Min Hwangbo"
date: "7/13/2020"
output:
  html_document:
    preserve_yaml: true
    toc: true
    toc_float: true
    keep_md: true
published: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* Inspiration: [本橋智光, 2019](http://www.yes24.com/24/AuthorFile/Author/278360)
* Data Source: [ECEAP Sites](https://data.wa.gov/Education/DCYF-ECEAP-Sites/8ydb-ddzd) @ Data.WA.GOV

# Step 1: Load packages
```{r, warning = F, message = F, results = "hide"}
# Install packages first using install.packages("pacakgename") on your console!

library("dplyr") # Data reshaping package
library("tidyr") # Data reshaping package
library("readr") # Reading a csv file
```

# Step 2: Preparation

## Load your file & Quality Check
Recommended to load `csv` file when possible using `readr` package and `read_csv` function.

```{r}
Original <- read_csv("ECEAPSites_DCYF_070120.csv")
# Original file

summary(Original, 5) # Summarize what this file is about
head(Original, 5) # First 5 records
tail(Original, 5) # Last 5 records
ls(Original) # Vars name

# 391 obs & 23 vars?
## But if you carefully check the original data from the website, it has the name of variables: Try View(Original) on your console
```

## Change Var Names

* First, we're going to copy information on the 2nd row to the 1st row as Var Names.
* Second, we're going to delete the duplicate record.
* Last, we will check the data quality/code accuracy.

```{r}
names(Original) <- Original[1,] # Copy 2nd row as var names
Revision <- Original[-1,] # Delete duplicate (1st) row

head(Revision, 5) # First 5 records
tail(Revision, 5) # Last 5 records
ls(Revision) # Vars name
# 390 obs & 23 vars
```

# Step 3: Data Extract

## Extract Certain Variables
What if we're only interested in extracting certain variables such as:

* Emergency Contact Name
* Site Name
* City
* Zip Code
* EAParticipation

Let's use `select` function!
```{r}
ECEAP1 <- Revision %>% 
  select("Emergency Contact Name", "Site Name", "City", "State", "Zip Code", "EAParticipation") %>% 
as.data.frame()
# 390 obs & 6 vars
```

Then... you probably remember we should at least check:

* First five records
* Last five records
* Vars name

```{r}
head(ECEAP1, 5) # First 5 records
tail(ECEAP1, 5) # Last 5 records
ls(ECEAP1, 5)
# Looks good!
```

## Renaming variable name
* Some of the variable names are way too long or have spaces in between.
* Let's change those names using `rename` function!
* "NewName" = `CurrentVarName`
```{r}
ECEAP2 <- ECEAP1 %>% rename("EATF" = `EAParticipation`, "EmergencyContact" = `Emergency Contact Name`, "SiteName" = `Site Name`, "ZipCode" = `Zip Code`) # For renaming dataframe column
# Don't forget to conduct your quality check!

head(ECEAP2, 5) # First 5 records
tail(ECEAP2, 5) # Last 5 records
ls(ECEAP2, 5) # Vars name
# Clear!
```

## Extract based on condition: `EATF = T/F`
* What if we're only interested in looking at ECEAP sites who are participating in Early Achievers?
* Let's use `filter` function!
```{r}
ECEAP3 <- ECEAP2 %>% 
  filter(EATF == "Yes")

# Time for quality check!
## View data set on your console! View(ECEAP3)
head(ECEAP3,5) # First five records
tail(ECEAP3,5) # Last five records
ls(ECEAP3) # Vars name

# Should have equal or smaller obs!
## 375 obs & 6 vars
### Clear!
```

# Step 4: Split data: `Extract`
Let's split name from `EmergencyContact` to 1) FirstName and 2) LastName
```{r}
ECEAP4 <- extract(ECEAP3, "EmergencyContact", c("FirstName", "LastName"), "([^ ]+) (.*)") # Still can't figure out this code - Retrieved from Stackflow

# You'll notice on the Environment that now we have 7 vars!
head(ECEAP4, 5)
tail(ECEAP4, 5)
ls(ECEAP4)
```

And we probably won't need TF var anymore for the final version if we think about adopting these codes for other project (i.e. CTM) 
```{r}
Final <- ECEAP4 %>% 
  select("FirstName", "LastName", "City", "State", "ZipCode") %>% 
as.data.frame()
# Should have 5 vars

head(Final, 5)
tail(Final, 5)
ls(Final)
```

# Step 5: Save as a CSV `write.csv`
* FileName_ProjectName_MMDDYY(DateCreated)
```{r}
write.csv(Final, "EASites_ECEAP_071220")
# Make sure to check your final csv data. 
## R will likely generate id on the first column.
```